---
title: "Collecting_Census_Inc_Data"
author: "Fiona Edwardson"
date: "5/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(sf)
library(tigris)
library(rmapshaper)
library(tmap)
library(areal)
library(leaflet)
library(dbplyr)
#install.packages("censusapi")
library(censusapi)
```

```{r putting in Census API key}

#census_api_key("MY_KEY", install = TRUE, overwrite = TRUE)

#Sys.getenv("CENSUS_API_KEY")

#readRenviron("~/.Renviron")
```

```{r confirm you're in the GISIII_Final_Project directory}

getwd()

```


# data: median household income

Data for all four cities

```{r pulling in the data}
chi_income2019 <- get_acs(geography = "tract", 
              year = 2019,
              variables = c(med_earn = "S2001_C01_002E"),
              state = "IL",
              county = "Cook",
              survey = "acs5",
              output = "wide",
              geometry = TRUE)

View(chi_income2019)


mpls_income2019 <- get_acs(geography = "tract", 
              year = 2019,
              variables = c(med_earn = "S2001_C01_002E"),
              state = "MN",
              county = "Hennepin",
              survey = "acs5",
              output = "wide",
              geometry = TRUE)

sd_income2019 <- get_acs(geography = "tract", 
              year = 2019,
              variables = c(med_earn = "S2001_C01_002E"),
              state = "CA",
              county = "San Diego",
              survey = "acs5",
              output = "wide",
              geometry = TRUE)


```


Now I need to pull in the city boundaries for each one:

```{r chicago boundaries}

chibounds = st_read("city_boundaries/Chicago_boundaries/geo_export_3d084884-b126-4daa-b908-1286d52a1e19.shp")

chibounds <- chibounds[0]
plot(chibounds[0])
#plot(chi_income2019)
chi_inc <- chi_income2019["med_earn"]
#print(chi_inc$med_earn)
plot(chi_inc)  ## this shows med income of all of Cook County income
glimpse(chi_inc)
```


```{r intersection}
st_crs(chi_inc)
st_crs(chibounds)

# setting NAD83 as the CRS
chibounds_CRS <- st_transform(chibounds, st_crs(chi_inc))
st_crs(chibounds_CRS)
chib_inc <- st_intersection(chi_inc, chibounds_CRS)

plot(chib_inc)
```


Now I just need to do this for the other cities: (it would also be cool to develop this into an algorithm...)

```{r mpls}

mpls_bounds <- st_read("city_boundaries/mpls_bounds/16cdbbfa-ad10-493c-afaf-52b61f2e76e42020329-1-180h9ap.whbo.shp")
mpls_b <- mpls_bounds[0]

mpls_inc <- mpls_income2019["med_earn"]
plot(mpls_b)

st_crs(mpls_b)
st_crs(mpls_inc)

# setting NAD83 as the CRS
mpls_b_CRS <- st_transform(mpls_b, st_crs(chi_inc))
st_crs(mpls_b_CRS)  # <- check
mplsb_inc <- st_intersection(mpls_inc, mpls_b_CRS)

plot(mplsb_inc)

```


```{r sd}

sd_bounds <- st_read("city_boundaries/san_diego_boundary_datasd/san_diego_boundary_datasd.shp")
sd_b <- sd_bounds[0]

sd_inc <- sd_income2019["med_earn"]
plot(sd_b)

st_crs(sd_b)
st_crs(sd_inc)

# setting NAD83 as the CRS
sd_b_CRS <- st_transform(sd_b, st_crs(sd_inc))
st_crs(sd_b_CRS)  # <- check
sdb_inc <- st_intersection(sd_inc, sd_b_CRS)

plot(sdb_inc)

```


## Constructing and Saving the final data

First I'll add a column identifying which city each census tract belongs to.

```{r adding city identifier column}

sdb_inc <- sdb_inc %>%
  mutate(city_name = "sd")

mplsb_inc <- mplsb_inc %>% mutate(city_name = "mpls")

chib_inc <- chib_inc %>% mutate(city_name = "chi")

```

It's occurred to me that the previous step wasn't strictly necessary, as it would be rather a pain to join all the tables together, and not necessary. So I'll save them as-is.


I need to do some final edits:
```{r exporting data}

getwd()  # always check! 

st_write(sdb_inc, "city_income_data/sd/sd_inc.shp")
st_write(mplsb_inc, "city_income_data/mpls/mpls_inc.shp")
st_write(chib_inc, "city_income_data/chi/chi_inc.shp")


```

.... Ok now I'm trying to do a basic clustering analysis with pysal, which means I think I want the data to be in .geojson formation.

```{r exporting data for python}

st_write(sdb_inc, "city_income_data/inc_data_geojson/sd/sd_inc.geojson")
st_write(mplsb_inc, "city_income_data/inc_data_geojson/mpls/mpls_inc.geojson")
st_write(chib_inc, "city_income_data/inc_data_geojson/chi/chi_inc.geojson")

```

Probably would have been a good idea to just use geojson files all along, but it's ok. (Maybe I would even be able to do a data package with the geojson files?)


