---
title: "Visualizing_Inc_Data"
author: "Fiona Edwardson"
date: "5/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We need some libraries for the visualizations!
```{r loading libraries...}

library(sf)
library(raster)
library(dplyr)
library(spData)
library(spDataLarge)
library(tmap)    
library(leaflet) 
library(ggplot2)

```


Let's make sure everything went as expected in the data wrangling process:

```{r pulling in chi income data}

c_inc <- st_read("city_income_data/chi/chi_inc.shp")
plot(c_inc)

# success!

```

Nice! Let's pull in the other files:

```{r pulling in mpls and sd income data}

sd_inc <- st_read("city_income_data/sd/sd_inc.shp")
plot(sd_inc)
glimpse(sd_inc)

m_inc <- st_read("city_income_data/mpls/mpls_inc.shp")
plot(m_inc)


```

Oh, it looks like I never ran the line that added the "sd" city name to the data set. If it's important, I can add it later. But worth noting!

Now I'll decide on how to format my choropleth maps. Here are the boundaries I'm thinking of:

- $ == city minimum wage 

chi == $15 / hr  == $31,200 per year
mpls == $15 / hr == $31,200 per year
sd == $ 15 / hr == $31,200 per year

I didn't actually know that! Neat.

- $ == can afford housing (avg 1br apt rent x 3)

chi = $2,080 == monthly income of $6,240 == annual income of $74,880
mpls = $1,621 == annual income of $58,356
sd = $2,756 == annual income of $99,216

Wow, these are drastically different values.


- $100,000 == more than six figures
- $1,000,000 == more than a million dollars

```{r inputting the correct choropleth boundaries}
# making a bigger boundary box to resolve title overlap issues
### note: pulled this code from the internet! https://stackoverflow.com/questions/60892033/how-do-you-position-the-title-and-legend-in-tmap

bbox_new <- st_bbox(c_inc) # current bounding box

xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
yrange <- bbox_new$ymax - bbox_new$ymin # range of y values

#bbox_new[1] <- bbox_new[1] - (0.25 * xrange) # xmin - left
#bbox_new[3] <- bbox_new[3] + (0.25 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.25 * yrange) # ymin - bottom
bbox_new[4] <- bbox_new[4] + (0.1 * yrange) # ymax - top

bbox_new <- bbox_new %>%  
  st_as_sfc() # turns the bbox into a sf polygon


## Chicago map
chibreaks = c(12600, 31200, 74880, 100000, 500000)
tm_shape(c_inc, bbox = bbox_new) + tm_polygons(col = "med_earn", breaks = chibreaks, title = expression("Median Income")) + tm_layout(bg = "grey99", legend.position = c("left", "bottom"), legend.title.size = 0.8,legend.text.size = 0.5, title = "Individual Median Income in Chicago") 

#tm_fill(legend_title = expression("Indvl Median Income"))

```

```{r mpls map}
bbox_new <- st_bbox(m_inc) # current bounding box

xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
yrange <- bbox_new$ymax - bbox_new$ymin # range of y values

bbox_new[1] <- bbox_new[1] - (0.6 * xrange) # xmin - left
#bbox_new[3] <- bbox_new[3] + (0.25 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.25 * yrange) # ymin - bottom
bbox_new[4] <- bbox_new[4] + (0.1 * yrange) # ymax - top

bbox_new <- bbox_new %>%  
  st_as_sfc() # turns the bbox into a sf polygon


mplsbreaks = c(12600, 31200, 58356, 100000, 500000)
tm_shape(m_inc, bbox = bbox_new) + tm_polygons(col = "med_earn", breaks = mplsbreaks, title = expression("Median Income")) + tm_layout(bg = "grey99", legend.position = c("left", "bottom"), legend.title.size = 0.8,legend.text.size = 0.5, title = "Individual Median Income in Minneapolis") 

#tm_fill(legend_title = expression("Indvl Median Income"))



```

```{r sd map}

bbox_new <- st_bbox(sd_inc) # current bounding box

xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
yrange <- bbox_new$ymax - bbox_new$ymin # range of y values

#bbox_new[1] <- bbox_new[1] - (0.1 * xrange) # xmin - left
#bbox_new[3] <- bbox_new[3] + (0.25 * xrange) # xmax - right
# bbox_new[2] <- bbox_new[2] - (0.25 * yrange) # ymin - bottom
bbox_new[4] <- bbox_new[4] + (0.15 * yrange) # ymax - top

bbox_new <- bbox_new %>%  
  st_as_sfc() # turns the bbox into a sf polygon


sdbreaks = c(12600, 31200, 99216, 100000, 500000)
tm_shape(sd_inc, bbox = bbox_new) + tm_polygons(col = "med_earn", breaks = sdbreaks, title = expression("Median Income")) + tm_layout(bg = "grey99", legend.position = c("right", "bottom"), legend.title.size = 0.8,legend.text.size = 0.5, title = "Individual Median Income in San Diego") 

#tm_fill(legend_title = expression("Indvl Median Income"))



```

